---
title: "Data generation"
author: "Koen Hufkens"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Data generation}
  %\VignetteEngine{knitr::rmarkdown}
  %\usepackage[utf8]{inputenc}
---

```{r include=FALSE}
library(dplyr)
library(FluxDataKit)
library(FluxnetLSM)
library(FluxnetEO)
```

# Compiling data

As mentioned in the introduction (once all required data are downloaded) the
`FluxDataKit` package ensures the proper compilation of `rsofun` driver data.
Although we will distribute a finished dataset the below instructions allow you
to recreate these data for a particular site (or sites).

To generate consistent data from FLUXNET formatted sources we need a site list
with some additional meta-data. A site list is generated using the script as 
described in the 'data coverage' vignette. We refer to this vignette to compile
the list of sites which can be processed.

Once a site list has been compiled you can use it (and all the other input data)
to generated either land surface model or `rsofun` compatible datasets. Here,
the former is used as a precursor to the latter.

## LSM formatting

By default land surface model compatible data is generated using the FluxnetLSM
package. Retaining only this data can be done by specifying the `format`
parameter, and setting it to "lsm". This routine will only save the netcdf
intermediates that are otherwise used for formatting p-model compatible data
and will not include any other ancillary data.

```{r eval = FALSE}
# load the sites to process
sites <- readRDS("data/flux_data_kit_site-info.rds") %>%
  filter(
    sitename == "BE-Vie"
  )

# process the data
fdk_process_lsm(
  sites,
  out_path = "/your/output/directory/",
  format = "lsm",
  modis_path = "data/modis/",
  overwrite = TRUE
  )
```

## p-model (rsofun) formatting

By default the format parameter is set to "lsm", providing land surface model
netcdf files as output. You can specify "fluxnet" to convert the data to
`rsofun` compatible FLXUNET output.

In addition, MODIS data can be merged from the FluxnetEO dataset using
the R package [with the same name](https://github.com/computationales/FluxnetEO). 
The latter ensures that rsofun driver (and target) data are ammended with MODIS
data for, among others, machine learning projects.

```{r eval = FALSE}
fdk_process_lsm(
  sites,
  out_path = "/your/output/directory/",
  format = "p-model",
  overwrite = TRUE
  )
```
