---
title: "Data use"
author: "Koen Hufkens"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Data use}
  %\VignetteEngine{knitr::rmarkdown}
  %\usepackage[utf8]{inputenc}
---

# p-model use

The (main) data is `rsofun` p-model compatible data. This means that the data
can be used to run the model (and optimizations) as is. We'll include a small
worked example below but refer to the `rsofun` 
[documentation](https://computationales.github.io/rsofun/articles/pmodel_use.html)
for more details.

Take note that not for all sites there is full coverage of the GPP data, these values
are then reported as NA. NA values should be handled per ususal in R and are safely
ignored during model optimization in `rsofun`. For more details on model optimization
we refer to the [vignette in the rsofun package](https://computationales.github.io/rsofun/articles/pmodel_use.html#calibrating-model-parameters).

```{r include=FALSE}
#library(Matrix) # hack because tidyverse bug ffs
library(dplyr)
library(rsofun)

# read in the data, only retain the first two sites for this demo
df <- readRDS("../data/p_model_drivers/site_based_drivers.rds")[1,]
modis <- readRDS("../data/modis_gee_data.rds") %>%
  rename(
    'modis' = 'data'
  )
```


```{r eval=FALSE}
library(rsofun)
library(dplyr)

# read in the data
df <- readRDS("data/p_model_drivers/site_based_drivers.rds")[1,]
```

```{r}
# optimized parameters from previous
# work
params_modl <- list(
    kphio           = 0.09423773,
    soilm_par_a     = 0.33349283,
    soilm_par_b     = 1.45602286,
    tau_acclim_tempstress = 10,
    par_shape_tempstress  = 0.0
  )
```


```{r}
# run the model for these parameters
output <- rsofun::runread_pmodel_f(
  drivers = df,
  par = params_modl,
  makecheck = FALSE
  )

print(output)
```

# MODIS data

By default the provided MODIS data in this repository is not merged with the
p-model drivers. To combine both datasets you can simply merge them by
sitename.

```{r eval = FALSE}
# load modis data and rename the data column
modis <- readRDS("data/modis_gee_data.rds") %>%
  rename(
    'modis' = 'data'
  )
```

Then do a simple left join on the available p-model data.

```{r}
# join datas
df <- left_join(df, modis, by = "sitename")
print(df)
```
All MODIS data is provided in a long format, so a conversion to a wide format
sorted by date is required before merging it with the forcing data as provided
for p-model model exercises.
